---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Import libraries
```{r echo=TRUE, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
library(stringr)
```

## Loading and preprocessing the data
```{r echo=TRUE}
# Read csv file and add factor for weekday/weekend. Convert steps to numeric.
DAT.FULL <- read.csv("activity.csv", stringsAsFactors = FALSE) %>%
        mutate(day_type = as.factor(ifelse(day(date) < 6, "weekday", "weekend")),
               steps = as.numeric(steps))
# Make a copy with NA values removed
DAT <- na.omit(DAT.FULL)
```


## What is mean total number of steps taken per day?
**Histogram of total number of step taken per day**
```{r echo=TRUE}
steps_per_day = DAT %>% group_by(date) %>% summarize(Total.Steps = sum(steps), Median=median(steps), Mean=mean(steps))

ggplot(data = steps_per_day, aes(x=Total.Steps)) + geom_histogram(bins=50) + scale_y_discrete("Number of Days")
```

**Total, median, and mean steps taken per day**
```{r echo=TRUE}
print(tbl_df(steps_per_day), n=100)
```

## What is the average daily activity pattern?
**Average steps by Interval**
```{r echo=TRUE}
plot_dat <- DAT %>% group_by(interval) %>% summarize(Avg.steps=mean(steps))
ggplot(plot_dat, aes(x=interval, y=Avg.steps)) + geom_line()
```
**Highest interval**
```{r echo=TRUE}
plot_dat %>% arrange(desc(Avg.steps)) %>% head(n=1)
```

## Inputing missing values
**Number of missing values**
```{r echo=TRUE}
sum(is.na(DAT.FULL))
```

**Fill in missing values**
```{r echo=TRUE}
# Compute average steps for the given interval, ignoring NA values
avg.steps <- function(myInterval) { 
        mean(na.omit(filter(DAT.FULL, interval==myInterval)$steps))
}
# Fill in NA values with average value
DAT.ENRICHED <- DAT.FULL %>% mutate(steps = ifelse(is.na(steps), avg.steps(interval), steps))
```

**Histogram of total number of step taken per day**
```{r echo=TRUE}
steps_per_day = DAT.ENRICHED %>% group_by(date) %>% summarize(Total.Steps = sum(steps), Median=median(steps), Mean=mean(steps))

ggplot(data = steps_per_day, aes(x=Total.Steps)) + geom_histogram(bins=50) + scale_y_discrete("Number of Days")
```

**Total, median, and mean steps taken per day**
```{r echo=TRUE}
print(tbl_df(steps_per_day), n=100)
```

### Comparison of enriched data set to unenriched
**Q:** Do these values differ from the estimates from the first part of the assignment?

**A:** Yes, the histogram has changed, and several days have different median and mean values.

**Q:** What is the impact of imputing missing data on the estimates of the total daily number of steps?

**A:** In general the number of total daily steps increases on days where steps values were NA.

## Are there differences in activity patterns between weekdays and weekends?
**Comparison of weekday vs weekend**
```{r echo=TRUE}
plot_dat <- DAT %>% group_by(interval, day_type) %>% summarize(Avg.steps=mean(steps))
ggplot(plot_dat, aes(x=interval, y=Avg.steps)) + geom_line() + facet_grid(day_type ~ .)
```